---
title: "Story Board Visualization"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    orientation: columns
    vertical_layout: fill
    theme: united
---

```{r setup, include=FALSE}
# Loading required libraries
library(flexdashboard)
library(readr)
library(ggplot2)
library(plotly)
```

Analysis {.storyboard}
=========================================

```{r,warning=FALSE,message=FALSE,echo=FALSE}
#Setting up the data object due to issues regarding parsing of .csv file
cdt <- read_csv("C:/Users/demon/Desktop/custodial-deaths.csv")
cd<-cdt[,1:12]

newnames<-c("Department.Name", "Custody.Type", "Custody.Type.Facility","First.Name","Middle.Name","Last.Name","Ethnicity","Sex","Age","Death.Date","Manner.Death","Charges.Status")
colnames(cd)<-newnames

# Adding aditional columns
cd$AgeGroup<-""
cd$AgeGroup<-ifelse(cd$Age<=0,NA,ifelse(cd$Age<=25,"1-25",ifelse(cd$Age<=50,"26-50",ifelse(cd$Age<=75,"51-75",ifelse(cd$Age<=100,"76-100",ifelse(cd$Age<=125,"101-125","Over 125"))))))

# Modifying to work with dates effectively
date<-strptime(cd$Death.Date, "%m/%d/%Y %H:%M")
cd$Death.Date<-date
```

### The dataset contains a larger proportion of male inmates as compared to female inmates  
  
```{r,warning=FALSE,message=FALSE,echo=FALSE}
p1<-ggplot(data=cd, aes(x=Sex,fill=AgeGroup))+geom_bar(width=1)+coord_polar(theta="y")+guides(fill= guide_legend(title = "Age Group"))+theme_light()
p1
```  
  
***  
  
  Another observation we can make is that the male inmate distribution represents a wider range of age groups as compared to the female inmates which have a higher concentration in the 26-50 age range, followed by 51-75.  
  
### The Age at the time of death across ethnicities shows considerable variation, especially for the male population  

```{r,warning=FALSE,message=FALSE,echo=FALSE,fig.align='center',fig.height=4,fig.width=7}
p2<-ggplot(data = cd, aes(x=Sex, y=Age, colour=Ethnicity, title("Age wise distribution")))+geom_boxplot()+coord_flip()+scale_color_brewer(palette = "Paired")+theme_dark()
p2
```  

***  
  
  + Another detail we can notice is that there are a few cases in which the age at the time of death was not recorded, specifically in the cases of __African-American__, __Anglo__ and __Hispanic__. There is also one case in the category of unknown ethnicity.
  + Another interesting observation that sticks out is that the inmates of __Native Hawaiian/Pacific Islander__ ethnicity shows a small group of 26-50 year olds who have passed away over the years.
  + We can also observe that in the case of female inmates, the only ethnicities observed were __African-American__, __Anglo__ and __Hispanic__.  
  
### The manner of death of inmates who were __Male__ shows considerable variation and a number of outlier cases. For __Female__ inmates, there are fewer outlier cases.

```{r,warning=FALSE,message=FALSE,echo=FALSE,fig.align='center',fig.height=4,fig.width=7}
p3<-ggplot(data = cd, aes(x=Sex,y=Age,fill=Manner.Death))+geom_boxplot()+coord_flip()+scale_fill_brewer()+theme_dark()
p3
```  

***  
  
  + Another observation that can be made is that there are fewer cases of __Female__ inmates who died due to __Accidental Injury cause by Others__, whose age were less than 25 years.  
  + Additionally, there are very few cases of __Female__ inmates whose cause of death were classified as __Other Homicide__, who were in their early to mid 40's.  
  + Another observation we can make is that there are significant number of outliers in the case of __Male__ inmates whose cause of death was classififed as __Natural Causes/Illness__, followed by those whose death was calssified as __Suicide__.
  
### On a yearly basis, one can observe that __Penitantiaries__ and __County Jails__ account for most of the deaths classified as __Natural Causes/Illness__  

```{r,warning=FALSE,message=FALSE,echo=FALSE,fig.align='center',fig.height=6,fig.width=11}
p4<-ggplot(data=cd, aes(x=factor(format(Death.Date, '%Y')),fill=Manner.Death))+geom_bar(position = "dodge")+scale_fill_brewer()+guides(fill= guide_legend(title = "Manner of Death"))+xlab("Year")+facet_wrap(~Custody.Type)+theme_dark()
p4
```  

***  
  
  + Most of the deaths that contribute to those classified as __Suicide__ also took place mostly at __Penitentiaries__ and __County Jails__.  
  + Another aspect we can observe is that the number of deaths in __Police Custody__ that are classified as __Justfiable Homicide__ are quite high, even when compared to other institutions.
  + Again, we see that __Muncipal Jails__ account for a few deaths comparatively.  
  
### For deaths in __Police Custody__ that have been deemed as __Justfiable Homicide__, a larger number of such deaths took place in 2013   

```{r,warning=FALSE,message=FALSE,echo=FALSE,fig.align='center',fig.height=4,fig.width=7}
# Subsetting data for construction of next visualization 
cdt1<-cd[cd$Custody.Type=="Police Custody (pre-booking)" & cd$Manner.Death=="Justifiable Homicide",]
  
# Plotting data
p5<-ggplot(data=cdt1, aes(x=factor(format(Death.Date, '%Y')),fill=Ethnicity))+geom_bar(position="dodge")+scale_fill_brewer(palette = "Paired")+xlab("Year")+theme_dark()
p5
```  

***  
  
  + We can also see that  initially, such deaths would happen more often to inmates of __Hispanic__ ethnicity but later years saw more of a blance between __Anglo__ and __Hispanic__ ethnicities.
  + It is interesting to see that the inmates whose ethnicity was classified as __Other/Don't Know__ have dates of death only in the year 2015.
  + Another observation that can be made is that the number of deaths of __Anglo__ ethnicity inmates, whose death was classified as __Justifiable Homicide__, seems to have spiked to the highest amount in 2013 (which is the most number of such deaths across ethnicities).  
  
### Generally, deaths in __Police Custody__ which are classified as __Justifiable Homicide__ tends to happen to inmates in an age range of 20-40 across ethnicities  

```{r,warning=FALSE,message=FALSE,echo=FALSE,fig.align='center',fig.height=4,fig.width=7}
p6<-ggplot(data= cdt1, aes(x=Sex,y=Age,fill=Ethnicity))+geom_boxplot()+coord_flip()+scale_fill_brewer(palette = "Paired")+theme_dark()
p6
```  

***  
  
  + Another observation that can be made is that there a number of outliers in the case of __Male__ inmates of __African-American__ and __Hispanic__ descent who have died in this way, whose age was above 50.  
  + There were also comparatiely less __Female__ inmates of __Hispanic__ descent who died this way (Usually of 50 years of age approximately), and also comparatively less __Male__ inmates whose ethnicity was unknown.  
  
### The Department where such deaths occured the most is the __Houston Police Department__ with __97__ of such occurences  

```{r,warning=FALSE,message=FALSE,echo=FALSE,fig.align='center',fig.height=4,fig.width=7}
# Calculating top 5 departments
cdfreq<-data.frame(table(cdt1$Department.Name))
cdfreq1<-cdfreq[order(-cdfreq$Freq),]
temp<-cdfreq1[1:5,]
disp<-cdt1[(cdt1$Department.Name=="Houston Police Dept.")|(cdt1$Department.Name=="San Antonio Police Dept.")|(cdt1$Department.Name=="Dallas Police Dept.")|(cdt1$Department.Name=="Texas Department Of Public Safety")|(cdt1$Department.Name=="Fort Worth Police Dept."),]

# Plotting an interactive graph
p7<-ggplot(data=disp, aes(x=Department.Name, fill=Department.Name))+geom_bar()+theme_light()+theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+xlab("Department Name")
(gg<-ggplotly(p7))
```  

***  
  
  + There are two departments with the same amount of such occurences, namely __Dallas Police Department__ and __San Antonio Police Department__ with __61__ of such occurences.
  + While not shown here, there are approximately 117 more departments where such occurences have taken place over the years, but these are usually of the order of 1 to 3.  
  
Details
=========================================

Column
-----------------------------------------

### Team Members:  
  + Ashish Syal
  + Dean D'souza
  + SoujanyaPrakash Bulagannawar
  
  