---
title: "ANLY 512 - Problem Set 2"
subtitle: "Visualizing Movie's"
author: "Dean D'souza"
date: "`r Sys.Date()`"
output: html_document
---


## Objectives

The objectives of this problem set is to gain experience working with data visualization as a tool for understanding unfamiliar data. To do this I want you to submit four *vis* that reveal something interesting about either the included `movies.csv` dataset.  For each *vis*, include the code you used to create the plot (`echo=TRUE`), and a one paragraph description of what you found interesting.  In most analyses the first plot you try will not be the most informative and you will need to make multiple improvements to get a polished plot. To show this process, for one of your 4 *vis*, provide 3-5 iterations that lead you to the final plot.  

Finally, upon completion name your final output `.html` file as: `YourName_PS2_ANLY512-50_SU2016.html` and upload it to the "Problem Set 2"" assignment on Moodle.


## My Example

Below is my example for a multi-part question.  I have also included an example of loading the data based on requests in last weeks class.  I will provide the `movie_data.csv` check Moodle.  Add your response in a similar fashion and submit following the directions above.

### Data Import & Cleanup
```{r, echo=TRUE}
library(dplyr)  # for easier filtering
library(ggplot2)

dat <- read.csv("C:/Users/Dean/Desktop/DataViz/movie_data.csv")  # import the csv

str(dat)    # Look at the structure of the data frame

# Clean up some variables
dat$Domestic <- as.numeric(dat$Domestic)   # Convert these columns to numeric
dat$Worldwide <- as.numeric(dat$Worldwide)

dat2 <- filter(dat, Title.type == "Feature Film")
```


### Multi-step Visualization

Look at the distribution of earnines by release year.
```{r, echo=TRUE}
ggplot(dat2, aes(Year, Worldwide)) +
  geom_bar(stat = "identity")
```

Lots of empty year earlier on, I will try just domestic earnings.
```{r, echo=TRUE}
ggplot(dat2, aes(Year, Domestic)) +
  geom_bar(stat = "identity")
```

No similar patterns.  Looks like missing data early in the time series.

```{r, echo=TRUE, fig.width=8.5}
dat3 <- filter(dat2, Year >= 1990)

ggplot(dat3, aes(factor(Year), Domestic)) +
  geom_bar(stat = "identity") +
  ggtitle("Annual Domestic Movie Earnings") +
  ylab("Domestic Earnings (USD)") +
  xlab("Year")
```

### Summary

There are several intersting details about this dataset I have identified.  First there is excessive missing data points early in the time-series.  A focus on data post-1990 is much more interesting.  Second, the data seems to have inconsistent `Year` data because I have found abberrant year designations.  In an analysis I would have to go back and ensure data is being aggregated to the year.  Finally, and most interestingly domestic movie earnings have been rather inconsistent over the last twenty year.  

## Visualization 1

For the first visualization we look at the Runtime in minutes for the entire data set.

### Multi-step Visualization

```{r}
# A bit of additional clean up

dat$Title.type[dat$Title.type=="Mini-series"]<-"Mini-Series"

# Plotting the visualization

gp4<-ggplot(dat, aes(x=Runtime..mins.))
gp4+geom_bar(na.rm = TRUE)+xlab("Runtime(mins)")

```
  
    From the above plot we see that most of the cases in our data set have a runtime between 0 and 500 minutes.  
    We further scale the plot down to get a better look at the distribution.
    
```{r}
gp4+geom_bar(na.rm = TRUE)+scale_x_continuous(limits=c(50,200))+xlab("Runtime(mins)")
```
  
  While the above plot is useful, we can better summarize the data by using `geom_density()` function:
  
```{r}
gp4+geom_density(na.rm=TRUE)+scale_x_continuous(limits=c(0,200))+xlab("Runtime(mins)")
```

### Summary

  The plot helps us see that the majority of cases in the dataset have a runtime close to 60 minutes. It helps us to understand that most of the cases are short movies and that there are lesser examples of other types such as music videos, mini-series, etc.
  
## Visualization 2
  
  Here we take a look at the variation of IMDB ratings over the years while grouping based on the Title type.
  
```{r}
dat.1<-filter(dat, Year>=1970 ,na.rm=TRUE)
gp3<-ggplot(dat.1, aes(x= Year, y= IMDb.Rating, fill= Title.type))
gp3+geom_bar(stat="identity", position="stack")+xlab("Year")+ylab("Rating on IMDB")
```

  As we have already seen, most of the cases in the data set are of the "Feature Film" and "Mini-series" type. Due to the composition of the data many of the other types are seen as short bands on the chart.

## Visualization 3
  
  In the following visualization we try to see if there is a relationship between the IMDB ratings and Average Ratings of the various cases. We also divide this based on the Title.type for a better view of variation across this factor.

```{r}
gp1<-ggplot(dat, aes(x=IMDb.Rating, y=AVG.rating, colour=Title.type))
gp1+geom_point(na.rm=TRUE)+facet_wrap(~Title.type)+xlab("Ratings on IMDB")+ylab("Average Rating")
```
  
  Based on the plot we can see that overall it would seem that there is a linear relationship between the IMDB ratings and Average ratings over the various groups. However in the case of "Feature Film" and of "Documentaries", we see a lot of variation.

## Visualization 4
  
  The following visualization takes a look at the variation of Average Ratings over the years for the various groups of Title.type.

```{r}
#dat$Rotten.Tom.<-as.numeric(dat$Rotten.Tom.)

#  Trying the exclude certain Title.type so as to only view proper violin plots
dat.2<-filter(dat,!(grepl("Feature Film and Mini-series|Music video|TV Episode|TV Special|NA",Title.type)))

#dat.3<-filter(dat.1, Rotten.Tom.>0,na.rm=TRUE)

# plotting the violin plots by faceting over Title.type
gp2<-ggplot(dat.2, aes(x= Year , y= AVG.rating , fill= Title.type))
gp2+geom_violin(na.rm=TRUE)+facet_wrap(~Title.type,drop = TRUE)+xlab("Year")+ylab("Average Rating")
```
  
  The above plot gives us an idea about the distribution of Average ratings over the years fot the varuious Title.type. 
  
## Summary

  The above visualizations help us to get a better understanding of the data set and how the various cases are distributed. The ggplot2 package provides a number of useful visualization tools that can help with depicting information better as well and the dplyr package helps to better structure the data for the visualizations. This document can also be found on [GitHub](https://github.com/DeanDsouza25/ANLY_512-50_SU2016/tree/master/Assignment_2)
