---
title: "Assignment 4"
subtitle: "Exploratory Data Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

The objective of this assignment is to familiarize your self with the process, specifically the visual process of Exploratory Data Analysis (EDA).  I have assigned several readings to support this work and layout an overall procedure.  You should read **Zuur et al.** prior to completing this work.


## Directions

Using any data set in `R`, for instance the `titatic` data set found within the `library(datasets)` package as your analysis data.  Conduct an EDA on the data set of your choice.  Your analysis should follow the "key steps" discussed in **Zuur et al.** and in our lecture discussion.  

You analysis should be completed within an **Rmarkdown** document.  The assignment should show data import, and any data cleanup that you conduct.  The document should show the steps in your EDA including:

- $\geq$ 5 statistical graphics.  
- Each graphic should include plot titles, proper axis labels, and legend information, if neccesary.  
- Along with each plot include a short statement indicating what you learned from that plot.  (This could be anything from verifying a `sort()` or `filter()` call to a `mosaic()` plot showing sample size variability).  

The final document should be converted using the `knitr` package to an `.html` file and submitted via Moodle for evaluation.  Remember to properly name your file as: `YourName_ANLY512-50_SU2016.html`.

## Your EDA

```{r}
# Your import statement
```


```{r}
# Your first EDA step
```
